Supplementary Information -- Morphological convergence as on-line lexical analogy

### 2020-01-28 models/gcm/gcm.Rmd
in progress

```{r }
source('Github/RaczBecknerHayPierrehumbert2019/functions/grabfunctions.R')

source('Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')

lookup = read_csv('Github/RaczBecknerHayPierrehumbert2019/models/gcm/feature_distance_lookup_tidy.txt')

@@

# test words
test = baseline.verbs$disc
# training words: verbs in celex in moder classes
training1 = celex %>% 
  filter(!is.na(moder)) %>% 
  select(ahpa) %>% # using Albright and Hayes' transcription
  pull
# training words: other regular verbs (monsyllabic)
c2 = celex %>% 
  filter(
    category == 'regular',
    is.na(moder),
    !is.na(ahpa)
    ) %>% 
  mutate(
    nchar = nchar(ahpa),
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'),
    nvowels = map(vowels, ~ length(.))
  ) %>% 
  select(-vowels) %>% 
  unnest()

@@

# batch 1
distances1 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out.txt', col_names = F)
names(distances1) = 'feature.distance'

pairs1 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody.txt', '\t', col_names = F)
names(pairs1) = c('target', 'training')

@@

# batch 2
distances2 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out2.txt', col_names = F)
names(distances2) = 'feature.distance'

pairs2 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody2.txt', '\t', col_names = F)
names(pairs2) = c('target', 'training')

@@

# batch 3
distances3 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out3.txt', col_names = F)
names(distances3) = 'feature.distance'

pairs3 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody3.txt', '\t', col_names = F)
names(pairs3) = c('target', 'training')

@@

# batch 4
distances4 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out4.txt', col_names = F)
names(distances4) = 'feature.distance'

pairs4 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody4.txt', '\t', col_names = F)
names(pairs4) = c('target', 'training')

@@

celex2 = celex %>% 
  # filter(!(is.na(moder) & category == 'irregular')) %>%
  filter(!is.na(ahpa)) %>% # celex has more rows than albright & hayes celex. we want the latter.
  mutate( # we want monosyllabic
    nchar = nchar(ahpa),
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'),
    nvowels = map(vowels, ~ length(.))
  ) %>% 
  select(-vowels) %>% 
  unnest()

>>>>

```{r }
source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabfunctions.R')

source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')

lookup = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/models/gcm/feature_distance_lookup_tidy.txt')

@@

# test words
test = unique(baseline$disc)
# training words: verbs in celex in moder classes
training1 = celex %>% 
  filter(!is.na(moder)) %>% 
  select(ahpa) %>% # using Albright and Hayes' transcription
  pull
# training words: other regular verbs (monsyllabic)
c2 = celex %>% 
  filter(
    category == 'regular',
    is.na(moder),
    !is.na(ahpa)
    ) %>% 
  mutate(
    nchar = nchar(ahpa),
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'),
    nvowels = map(vowels, ~ length(.))
  ) %>% 
  select(-vowels) %>% 
  unnest(cols = c(nvowels))

@@

# batch 1
distances1 = read_csv('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/out.txt', col_names = F)
names(distances1) = 'feature.distance'

pairs1 = read_delim('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody.txt', '\t', col_names = F)
names(pairs1) = c('target', 'training')

@@

# batch 2
distances2 = read_csv('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/out2.txt', col_names = F)
names(distances2) = 'feature.distance'

pairs2 = read_delim('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody2.txt', '\t', col_names = F)
names(pairs2) = c('target', 'training')

@@

# batch 3
distances3 = read_csv('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/out3.txt', col_names = F)
names(distances3) = 'feature.distance'

pairs3 = read_delim('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody3.txt', '\t', col_names = F)
names(pairs3) = c('target', 'training')

@@

# batch 4
distances4 = read_csv('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/out4.txt', col_names = F)
names(distances4) = 'feature.distance'

pairs4 = read_delim('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody4.txt', '\t', col_names = F)
names(pairs4) = c('target', 'training')

@@

celex2 = celex %>% 
  # filter(!(is.na(moder) & category == 'irregular')) %>%
  filter(!is.na(ahpa)) %>% # celex has more rows than albright & hayes celex. we want the latter.
  mutate( # we want monosyllabic
    nchar = nchar(ahpa),
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'),
    nvowels = map(vowels, ~ length(.))
  ) %>% 
  select(-vowels) %>% 
  unnest(cols = c(nvowels))

### 2020-01-28 models/paper/convergence_paper_models.Rmd
in progress

@@

```{r }
baseline %>% group_by(word) %>% summarise(regular = mean(regular)) %>% filter(word == 'splive') 
```

@@

```{r }
espdata %>% select(participant_id,lex_typicality,reg_rate) %>% unique %>% group_by(lex_typicality,reg_rate) %>% summarise(n = n()) 
```

@@

baseline %>% select(word,r.baseline_gcm_features,baseline_gcm_edits) %>% filter(word == 'splive') %>% unique
baseline %>% select(word,r.baseline_mgl_features,baseline_mgl_edits) %>% filter(word == 'splive') %>% unique

@@

```{r }
# don't want to load Hmisc, messes with select()
with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .581

with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) # .586
```

@@

save(m3, file = '../../models/paper/m3.rda')

@@

save(m4, file = '../../models/paper/m4.rda')
```

@@

load('../../models/paper/m4.rda')
load('../../models/paper/m3.rda')

summary(m3)
#                         Estimate Std. Error z value Pr(>|z|)    
# (Intercept)             -0.44918    0.11202  -4.010 6.08e-05 ***
# c.baseline_mgl_features  1.27846    0.07525  16.989  < 2e-16 ***
# gcm_residual             1.37767    0.17171   8.023 1.03e-15 ***
summary(m4)
#                         Estimate Std. Error z value Pr(>|z|)    
# (Intercept)             -1.12780    0.14259  -7.909 2.58e-15 ***
# c.baseline_gcm_features  2.29014    0.15126  15.140  < 2e-16 ***
# mgl_residual             0.95005    0.08549  11.113  < 2e-16 ***
```

@@

with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .58
with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) # .59

with(test2, Hmisc::somers2(baseline_gcm_features,resp_post_reg)['C']) # .6
with(test2, Hmisc::somers2(individual_gcm_features,resp_post_reg)['C']) # .68

with(test2, Hmisc::somers2(baseline_mgl_features,resp_post_reg)['C']) # .61
with(test2, Hmisc::somers2(individual_mgl_features,resp_post_reg)['C']) # .55
```

@@

```{r eval=F}
test2 = test2 %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features),
    c.individual_gcm_features = scales::rescale(individual_gcm_features),
    c.individual_mgl_features = scales::rescale(individual_mgl_features),
  )

@@

m5 <- glmer(resp_post_reg ~ c.baseline_gcm_features + gcm_wf_score_diff + pre_reg_av + (1 + c.baseline_gcm_features + gcm_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m5, file = '../../models/paper/m5.rda')

m6 <- glmer(resp_post_reg ~ c.baseline_mgl_features + mgl_wf_score_diff + pre_reg_av + (1 + c.baseline_mgl_features + mgl_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m6, file = '../../models/paper/m6.rda')

# Third

m7 <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m7, file = '../../models/paper/m7.rda')

@@

```{r }
load('../../models/paper/m5.rda')
load('../../models/paper/m6.rda')
load('../../models/paper/m7.rda')

summary(m5)
#                  Estimate Std. Error z value Pr(>|z|)    
# (Intercept)              -5.5677     0.3724 -14.949  < 2e-16 ***
# c.baseline_gcm_features   4.3687     0.4249  10.281  < 2e-16 ***
# gcm_wf_score_diff         2.6117     0.4696   5.562 2.67e-08 ***
# pre_reg_av                5.2887     0.6022   8.782  < 2e-16 ***
summary(m6)
#                        Estimate Std. Error z value Pr(>|z|)    
# (Intercept)              -5.0641     0.3374 -15.010   <2e-16 ***
# c.baseline_mgl_features   2.0335     0.2079   9.782   <2e-16 ***
# mgl_wf_score_diff         0.2085     0.1723   1.210    0.226    
# pre_reg_av                7.4492     0.5996  12.425   <2e-16 ***
summary(m7)
#                      Estimate Std. Error z value Pr(>|z|)    
# (Intercept)                -5.6326     0.3364 -16.743  < 2e-16 ***
# c.individual_gcm_features   2.7857     0.3776   7.378 1.61e-13 ***
# c.baseline_mgl_features     1.5998     0.1709   9.361  < 2e-16 ***
# pre_reg_av                  5.4374     0.5842   9.307  < 2e-16 ***
summary(m7)$coef %>% round(3) %>% xtable
# vif.mer(m7)
summary(rePCA(m7))

@@

```{r eval=F}
load('esp_fit.rda')
formula(fit)
MuMIn::r.squaredGLMM(fit) # 0.27 0.45
MuMIn::r.squaredGLMM(m7) # 0.22 0.47 
test2$m7.pred = predict(m7, type = 'response')
with(test2, Hmisc::somers2(m7.pred, resp_post_reg)) # .84
```

@@

t2 = test2 %>%
  mutate(lex_typicality2 = 
           case_when(
             lex_typicality == 'atypical' ~ 'reversed',
             lex_typicality == 'typical' ~ 'typical',
             lex_typicality == 'random' ~ 'random',
           ),
         lex_typicality2 = factor(lex_typicality2, levels = c('typical', 'reversed', 'random'))
         ) %>% 
  group_by(
    word, 
    lex_typicality2
  ) %>% 
  summarise(
    post.reg.av = mean(resp_post_reg),
    individual_gcm_features = mean(individual_gcm_features),
    individual_mgl_features = mean(individual_mgl_features),
  ) %>% 
  ungroup() %>% 
  mutate(
    individual_gcm_features = scales::rescale(individual_gcm_features, to = c(0,1)),
    individual_mgl_features = scales::rescale(individual_mgl_features, to = c(0,1)),
    post.reg.av = scales::rescale(post.reg.av, to = c(0,1))
  )

@@

ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2, linetype = lex_typicality2)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  # scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline regularization') +
  ylab('verb ESP regularization') +
  guides(linetype = F) +
  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
# ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
ggsave('fig4_2019.pdf', width = 12.5, height = 4)

>>>>

espdata[espdata$lex_typicality == 'atypical',]$lex_typicality = 'reversed'

test2[test2$lex_typicality == 'atypical',]$lex_typicality = 'reversed'

@@

```{r }
baseline %>% 
  group_by(word) %>% 
  summarise(regular = mean(regular)) %>% 
  filter(word == 'fim') 
```

@@

```{r }
espdata %>% 
  select(participant_id,lex_typicality,reg_rate) %>% 
  unique %>% 
  group_by(lex_typicality,reg_rate) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = reg_rate, values_from = n)
```

@@

baseline %>% 
  select(word,r.baseline_gcm_features,baseline_gcm_edits) %>% 
  filter(word == 'splive') %>% 
  unique

baseline %>% 
  select(word,r.baseline_mgl_features,baseline_mgl_edits) %>% 
  filter(word == 'splive') %>% 
  unique

@@

```{r }
# don't want to load Hmisc, messes with select()
with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C'])

with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) 
```

@@

save(m3, file = 'models/paper/m3.rda')

@@

save(m4, file = 'models/paper/m4.rda')
```

@@

load('models/paper/m4.rda')
load('models/paper/m3.rda')

summary(m3)

summary(m4)

@@

# first method

with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) 
with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C'])

with(test2, Hmisc::somers2(baseline_gcm_features,resp_post_reg)['C'])
with(test2, Hmisc::somers2(individual_gcm_features,resp_post_reg)['C'])

with(test2, Hmisc::somers2(baseline_mgl_features,resp_post_reg)['C']) 
with(test2, Hmisc::somers2(individual_mgl_features,resp_post_reg)['C'])
```

@@

# second method

test2 = test2 %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features),
    c.individual_gcm_features = scales::rescale(individual_gcm_features),
    c.individual_mgl_features = scales::rescale(individual_mgl_features),
  )

@@

m5 <- glmer(resp_post_reg ~ c.baseline_gcm_features + gcm_wf_score_diff + pre_reg_av + (1 + c.baseline_gcm_features + gcm_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m5, file = 'models/paper/m5.rda')

m6 <- glmer(resp_post_reg ~ c.baseline_mgl_features + mgl_wf_score_diff + pre_reg_av + (1 + c.baseline_mgl_features + mgl_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m6, file = 'models/paper/m6.rda')

# third method

m7a <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + c.individual_mgl_features + c.baseline_gcm_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))

# summary(m7a)

m7 <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m7, file = 'models/paper/m7.rda')

@@

```{r }
load('models/paper/m5.rda')
load('models/paper/m6.rda')
load('models/paper/m7.rda')

summary(m5)
summary(m6)
summary(m7)
summary(m7)$coef %>% round(3) %>% xtable
# vif.mer(m7)
summary(rePCA(m7))

@@

```{r eval=F}
load('models/paper/esp_fit.rda')
formula(fit)
MuMIn::r.squaredGLMM(fit) # 0.27 0.45
MuMIn::r.squaredGLMM(m7) # 0.22 0.47 
test2$m7.pred = predict(m7, type = 'response')
with(test2, Hmisc::somers2(m7.pred, resp_post_reg)) 
```

@@

t2 = test2 %>%
  mutate(lex_typicality2 = 
           case_when(
             lex_typicality == 'atypical' ~ 'reversed',
             lex_typicality == 'reversed' ~ 'reversed',
             lex_typicality == 'typical' ~ 'typical',
             lex_typicality == 'random' ~ 'random',
           ),
         lex_typicality2 = factor(lex_typicality2, levels = c('typical', 'reversed', 'random'))
         ) %>% 
  group_by(
    word, 
    lex_typicality2
  ) %>% 
  summarise(
    post.reg.av = mean(resp_post_reg),
    individual_gcm_features = mean(individual_gcm_features),
    individual_mgl_features = mean(individual_mgl_features),
  ) %>% 
  ungroup() %>% 
  mutate(
    individual_gcm_features = scales::rescale(individual_gcm_features, to = c(0,1)),
    individual_mgl_features = scales::rescale(individual_mgl_features, to = c(0,1)),
    post.reg.av = scales::rescale(post.reg.av, to = c(0,1))
  )

@@

ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2, linetype = lex_typicality2)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  # scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline regularization') +
  ylab('verb ESP regularization') +
  guides(linetype = F) +
  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
# ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
ggsave('fig4_2020.pdf', width = 12.5, height = 4)

### 2020-01-28 supplementary_information/si.tex
in progress

\begin{document}
\SweaveOpts{concordance=TRUE}

<<header, echo=FALSE, eval=TRUE>>=

# libs
library(tidyverse)
library(xtable)

# data
source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')

@

@@

<<header, echo=FALSE, eval=FALSE>>=

# table 1
baseline %>% 
  group_by(word) %>% 
  summarise(mean = mean(regular)) %>% 
  arrange(-mean)

@

@@

<<celex_counts, echo=F, eval=T>>=

# celex(f>=10) has repetitions:
celex_count = celex$disc %>% unique %>% length

gcm_table = celex %>% # take celex
  filter(!is.na(ahpa)) %>% # remove words not in Albright and Hayes celex
  mutate(
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'), # match vowels in the albright and hayes transcription
    nvowels = map(vowels, ~ length(.)) # calculate how many vowels the verb has
  ) %>% 
  select(-vowels) %>% # we don't need vowels now
  unnest(cols = c(nvowels)) %>%  # unnest nvowels
  filter(word != 'forsake', nvowels == 1) %>% # drop any verb with more than 1 vowel also "forsake which survives this
  select(category, moder, disc) %>% # keep category, moder, disc columns (= reg/irreg, drove/sang etc, transcription)
  distinct() %>% # repetitions! 
  count(category, moder) %>% # count drove/sang etc across reg/irreg
  mutate( # if it has no category, it's a misc verb
    moder = replace_na(moder, 'other')
  ) %>% 
  pivot_wider( # stretch table
    names_from = category, values_from = n
  ) %>% 
  rename(
   `verb class` = moder,
    `irregular set` = irregular,
    `regular set` = regular
  )

reg_count = filter(gcm_table, `verb class` == 'other') %>% 
  pull(`regular set`)

gcm_table = gcm_table[1:4,] %>% 
  mutate(
    `miscellaneous regular verbs` = reg_count
  )

gcm_table = rbind(
  gcm_table,
  c('sum', 
    sum(gcm_table$`irregular set`),
    sum(gcm_table$`regular set`),
    unique(gcm_table$`miscellaneous regular verbs`)
    )
)

@

Our starting point for the training set, following \cite{albright2003rules}, is the list of verbs in the \textsc{Celex} corpus (\citealt{baayen1993celex}, based on \citealt{sinclair1987}) with a token frequency of 10 or above, encompassing \Sexpr{celex_count} forms. However, similarity requirements restrict the respective training sets. We use the {\em DISC} transcription in which each contrastive segment of English is represented by a unique character (\texttt{dr2v} equals \textipa{[draIv]}).

@@

<<gcmtable, echo=F, eval=T, results=tex>>=
print(xtable(gcm_table, digits=2, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"))

@

\subsection{Estimation}

@@

<<splive1, echo=F, eval=T>>=

irreg_set = gcm_table %>% 
  filter(`verb class` == 'DROVE') %>% 
  pull(`irregular set`) %>% 
  as.double

reg_set = gcm_table %>% 
  filter(`verb class` == 'DROVE') %>% 
  pull(`regular set`) %>% 
  as.double

splive_gcm = test2 %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features)
  ) %>% 
  filter(word == 'splive') %>% 
  pull(c.baseline_gcm_features) %>% 
  unique %>% 
  round(2)

@

%The nonce form {\em splive} belongs to the DROVE class in our model. The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}. It is compared to \Sexpr{reg_set + reg_count} regular verbs -- these are \Sexpr{reg_set} verbs that match the DROVE schema (e.g.\ {\em side, hive, line}) and \Sexpr{reg_count} miscellaneous verbs. It is also compared to \Sexpr{irreg_set} irregular verbs (e.g.\ {\em drive, stride, smite}) in this class. Overall, it is more similar to the regular set: its {\em regularity score} is \Sexpr{splive_gcm}.

@@

<<mgl_form_counts, eval=T, echo=F>>=

# in order to look up the number of words per rule type, we need to dig up the raw mgl outputs.
# we have a function that turns them into mgl regular weights. but we need to break up that functions part-way and print the output here.

# get the raw output of the baseline feature-distance mgl
mgl_baseline_rules_raw = read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
  mutate(
    form1 = str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
    form2 = str_replace_all(form2, '»', '')
  )

disc_category = test2 %>% # get words and categories from test2. match colnames with above
  distinct(disc, category) %>% 
  mutate(
    form1 = disc
  )

verb_handles = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')

disc_reg_past_irreg_past = verb_handles %>% # get reg and irreg transcriptions from verb handles. match colnames with above.
  select(base.form,regular.form,irregular.form) %>% 
  rename(
    form1 = base.form
  )

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% # join them up
  left_join(disc_category, by = "form1") %>% 
  left_join(disc_reg_past_irreg_past, by = "form1")

# at this point, we only have categories etc for verbs in our ESP game. but since the interesting rules are those that apply to these forms, we focus on them only. so 156 verbs in total.

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% 
  filter(
    !is.na(category)
  )

# mgl_count$form1 %>% unique %>% length

# mind you, not all of these verbs have both regular and irregular rules, as discussed in the text. this is handled by the function processing the mgl output -- we won't deal with it here, since it's not relevant

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%  # group rules per output: reg / irreg!
  mutate(
    rule.type = case_when(
      form2 == regular.form ~ 'regular',
      form2 == irregular.form ~ 'irregular'
    )
  )

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%
  filter(
!is.na(rule.type)
)

# mgl_count$form1 %>% unique %>% length

# this is where it gets tricky. you can't sum up the hits and scope, because these overlap. you need to bin related forms and expressions per category per rule type and count them.

# first, we split the list-like strings into character vectors

mgl_count = mgl_baseline_rules_raw %>% # clean-up
  mutate(
    `related forms` = str_replace_all(`related forms`, '»', ''),
    exceptions = str_replace_all(exceptions, '»', ''),
    related_list = strsplit(`related forms`, ', '), # split
    exception_list = strsplit(exceptions, ', ')
  )

mgl_count1 = mgl_count %>%
  select(category,rule.type,related_list) %>% 
  unnest(related_list) %>% 
  distinct() %>% 
  count(category, rule.type) %>% 
  mutate(
    type = 'related forms'
  )

mgl_count2 = mgl_count %>%
  select(category,rule.type,exception_list) %>% 
  unnest(exception_list) %>% 
  distinct() %>% 
  count(category, rule.type) %>% 
  mutate(
    type = 'exceptions'
  )

mgl_table = rbind(mgl_count1,mgl_count2) %>% 
  pivot_wider(
    names_from = type, values_from = n
  ) %>% 
  arrange(rule.type, category)

@

<<mgltable, echo=F, eval=T, results=tex>>=
print(xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"))

@

\subsection{Estimation}

@@

<<splive2, echo=F, eval=F>>=

# In the next par, I won't pipe the values into variables in the text, but you can look them up here:
splive_mgl_rules = mgl_baseline_rules_raw %>% 
  filter(
    form1 == 'spl2v'
  )
# regular
splive_mgl_rules[3,]$Pfeat
splive_mgl_rules[3,]$hits
splive_mgl_rules[3,]$scope
splive_mgl_rules[3,]$reliability
splive_mgl_rules[3,]$confidence

# irregular
splive_mgl_rules[1,]$Pfeat
splive_mgl_rules[1,]$hits
splive_mgl_rules[1,]$scope
splive_mgl_rules[1,]$`related forms`
splive_mgl_rules[1,]$exceptions
splive_mgl_rules[1,]$reliability
splive_mgl_rules[1,]$confidence

test2 %>% 
  mutate(
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features)
  ) %>% 
  filter(word == 'splive') %>% 
  select(baseline_mgl_features,c.baseline_mgl_features) %>% 
  distinct()

@

@@

<<mgl_rule_counter, eval=T, echo=F>>=

mgl_rule_count = mgl_baseline_rules_raw %>% # take raw rules
  filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
    !is.na(category),
    !is.na(rule.type)
  ) %>% 
  mutate( # make a rule column of the form A -> B / C
    rule = paste0(A, ' ->', B, ' / ', Pfeat)
  ) %>% 
  distinct(rule,category,rule.type) %>% # keep DISTINCT rules
  count(category, rule.type) %>%  # count them
  pivot_wider(names_from = rule.type, values_from = n) %>% # stretch table
  mutate( # make row sum
    sum = irregular + regular
  ) %>% 
  rename(
    `regular rule` = regular,
    `irregular rule` = irregular
  )

# glue on colsums
mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))

@

<<mgltable, echo=F, eval=T, results=tex>>=
print(xtable(mgl_table, digits=2, label="mglruleses", caption="The number of unique rules posited by the MGL in each verb class"))

@

Table \ref{mglruleseskept} shows the list of (irregular and regular) rules for one of the classes, the KEPT class. We choose this class as it provides a good illustration of the estimation process. An MGL rule describes an $A \rightarrow{} B$ {\em alternation} in an $XAY$ {\em environment}. Since the KEPT rules are all suffixing, the following environment ($Y$) is empty. For the preceding environment ($X$), we give a list of segments that are matched by the structural description if it applies to four segments or fewer. Otherwise, we give a phonological description. So, for instance, the first rule turns \textipa{[il]} to \textipa{[Elt]} following any obstruent. Note that in some cases, the MGL generates multiple rules for the same alternation (such as \textipa{ip} $\rightarrow{}$ \textipa{Ept}). As discussed in Section \ref{mgloutput}, in such instances, we choose the candidate rule with the highest adjusted confidence in order to calculate the MGL regularity score. (Fricatives and nasal consonants constitute a natural class in the segmental feature system used by \citealt{albright2003rules}.)

>>>>

\usepackage{Sweave}
\begin{document}
\input{si-concordance}

@@

@@

Our starting point for the training set, following \cite{albright2003rules}, is the list of verbs in the \textsc{Celex} corpus (\citealt{baayen1993celex}, based on \citealt{sinclair1987}) with a token frequency of 10 or above, encompassing 3156 forms. However, similarity requirements restrict the respective training sets. We use the {\em DISC} transcription in which each contrastive segment of English is represented by a unique character (\texttt{dr2v} equals \textipa{[draIv]}).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Mon Jan 27 16:35:17 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllll}
  \hline
 & verb class & irregular set & regular set & miscellaneous regular verbs \\ 
  \hline
1 & BURNT & 6 & 42 & 1218 \\ 
  2 & DROVE & 14 & 83 & 1218 \\ 
  3 & KEPT & 12 & 31 & 1218 \\ 
  4 & SANG & 8 & 13 & 1218 \\ 
  5 & sum & 40 & 169 & 1218 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, GCM training data} 
\label{gcmsetsize}
\end{table}
\subsection{Estimation}

@@

%The nonce form {\em splive} belongs to the DROVE class in our model. The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}. It is compared to 1301 regular verbs -- these are 83 verbs that match the DROVE schema (e.g.\ {\em side, hive, line}) and 1218 miscellaneous verbs. It is also compared to 14 irregular verbs (e.g.\ {\em drive, stride, smite}) in this class. Overall, it is more similar to the regular set: its {\em regularity score} is 0.57.

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Mon Jan 27 16:35:17 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllrr}
  \hline
 & category & rule.type & related forms & exceptions \\ 
  \hline
1 & burnt & irregular &  24 &  41 \\ 
  2 & kept & irregular &  10 &   9 \\ 
  3 & sang & irregular &  11 &  77 \\ 
  4 & burnt & regular &  38 &  21 \\ 
  5 & drove & regular &  29 &  27 \\ 
  6 & kept & regular &  67 &  41 \\ 
  7 & sang & regular &  47 &  38 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, MGL training data} 
\label{mglsetsize}
\end{table}
\subsection{Estimation}

@@

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Mon Jan 27 16:35:17 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllrr}
  \hline
 & category & rule.type & related forms & exceptions \\ 
  \hline
1 & burnt & irregular &  24 &  41 \\ 
  2 & kept & irregular &  10 &   9 \\ 
  3 & sang & irregular &  11 &  77 \\ 
  4 & burnt & regular &  38 &  21 \\ 
  5 & drove & regular &  29 &  27 \\ 
  6 & kept & regular &  67 &  41 \\ 
  7 & sang & regular &  47 &  38 \\ 
   \hline
\end{tabular}
\caption{The number of unique rules posited by the MGL in each verb class} 
\label{mglruleses}
\end{table}
Table \ref{mglruleseskept} shows the list of (irregular and regular) rules for one of the classes, the KEPT class. We choose this class as it provides a good illustration of the estimation process. An MGL rule describes an $A \rightarrow{} B$ {\em alternation} in an $XAY$ {\em environment}. Since the KEPT rules are all suffixing, the following environment ($Y$) is empty. For the preceding environment ($X$), we give a list of segments that are matched by the structural description if it applies to four segments or fewer. Otherwise, we give a phonological description. So, for instance, the first rule turns \textipa{[il]} to \textipa{[Elt]} following any obstruent. Note that in some cases, the MGL generates multiple rules for the same alternation (such as \textipa{ip} $\rightarrow{}$ \textipa{Ept}). As discussed in Section \ref{mgloutput}, in such instances, we choose the candidate rule with the highest adjusted confidence in order to calculate the MGL regularity score. (Fricatives and nasal consonants constitute a natural class in the segmental feature system used by \citealt{albright2003rules}.)
